---
title: "Isotope data reduction"
date: "`r Sys.Date()`"
  html_document:
    code_folding: show
    df_print: paged
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
editor_options:
  chunk_output_type: console
---

```{r setup, include = FALSE}
# global knitting options for code rendering
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>")

# global knitting options for automatic saving of all plots as .png and .pdf
knitr::opts_chunk$set(
  dev = c("png", "pdf"),
  dev.args = list(pdf = list(encoding = "WinAnsi", useDingbats = FALSE)),
  fig.keep = "all",
  fig.path = file.path("fig_output", paste0(gsub("\\.[Rr]md", "", knitr::current_input()), "_"))
)
```

# Load packages

```{r, message=FALSE, warning=FALSE}
library(tidyverse) # general data wrangling and plotting
library(isoreader) # reading the raw data files
library(isoprocessor) # processing the data
```

This analysis was run using [isoreader](http://isoreader.kopflab.org) version `r packageVersion("isoreader")` and [isoprocessor](http://isoprocessor.kopflab.org/) version `r packageVersion("isoprocessor")`. 

All code chunks that contain a critical step towards the final data (i.e. do more than visualization or a data summary) are marked with `(*)` in the header to make it easier to follow all key steps during interactive use. 

# Load data

## Read raw data files (*)

```{r, warning=FALSE}
# set file path(s) to data files, folders or rds collections 
# can be multiple folders or mix of folders and files, using example data set here
# (this is a reduced data set of mostly standards with just a few example samples)
data_path <- file.path("data", "201911_isotope_data.cf.rds")

# read files
iso_files_raw <- 
  # path to data files
  data_path %>% 
  # read data files in parallel for fast read
  iso_read_continuous_flow() %>%
  # filter out files with read errors (e.g. from aborted analysis)
  iso_filter_files_with_problems()
```


## Process file info & peak table (*)

```{r}
# process file information
iso_files <- iso_files_raw %>% 
  # rename key file info columns
  iso_rename_file_info(analysis = Analysis, id1 = `Identifier 1`, id2 = `Identifier 2`) %>% 
  # parse text info into numbers
  iso_parse_file_info(number = analysis) %>% 
  # process other file information that is specific to the naming conventions
  # of this particular sequence
  iso_mutate_file_info(
    # what is the type of each analysis?
    type = case_when(
      str_detect(id1, "[Zz]ero")      ~ "on_off",
      str_detect(id1, "H3")           ~ "H3_factor",
      str_detect(id1, "F8")           ~ "std",
      TRUE                            ~ "sample"
    ),
    # what was the concentration? (assuming Preparation = concentration or volume)
    concentration = 
      ifelse(type == "std",  
             str_extract(Preparation, "[0-9.]+ ?ng( per |/)uL") %>% 
               parse_number() %>% iso_double_with_units("ng/uL"),
             NA),
    # what folder are the data files in? (assuming folder = sequence)
    folder = basename(dirname(file_path))
  ) %>% 
  # focus only on the relevant file info, discarding the rest
  iso_select_file_info(
    folder, analysis, file_datetime, id1, type, concentration
  ) %>% 
  # add in additional sample metadata (could be any info)
  # note: this would typically be stored in / read from a csv or excel file
  iso_add_file_info(
    readxl::read_excel(file.path("data", "isotopes_metadata.xlsx")),
    join_by = "id1"
  )

# set peak table from vendor data table with default isodat template
iso_files <- iso_set_peak_table_from_isodat_vendor_data_table(iso_files) %>% 
  # convert units from mV to V for amplitudes and area
  iso_convert_peak_table_units(V = mV, Vs = mVs)

# focus on sample files
sample_files <- iso_filter_files(iso_files, type == "sample" & analysis > 5651)
```


## Show file information

```{r}
# display file information
iso_files %>% 
  iso_get_file_info() %>% select(-file_id, -folder) %>% 
  iso_make_units_explicit() %>% knitr::kable()
```

## Example chromatograms

```{r "example_chromatograms", fig.width=8, fig.height=8}
# plot the chromatograms
iso_plot_continuous_flow_data(
  # all sample files
  sample_files,
  # select data and aesthetics
  data = c(2), color = id1, panel = analysis,
  # zoom in on time interval
  time_interval = c(1000, 3000),
  # peak labels for all peaks > 2V
  peak_bounds = TRUE,
  peak_marker = FALSE,
  peak_label = iso_format(rt),
  #peak_label_size = 3,
  peak_label_filter = analysis == 5685 & amp2 > 1
) 
```

# Peak Mapping

```{r}
# this information is often maintained in a csv or Excel file instead
# but generated here from scratch for demonstration purposes
peak_maps <- 
  readxl::read_excel(file.path("data","isotopes_peak_maps.xlsx"))
peak_maps %>% knitr::kable(digits = 0)

# identify peaks
sample_files_w_ids <- sample_files  %>% 
  # map peaks
  iso_map_peaks(peak_maps, map_id = type)
```

## Chromatograms

```{r "sample_chromatograms", fig.width=8, fig.height=8}
# plot the chromatograms
iso_plot_continuous_flow_data(
  # all sample files
  sample_files_w_ids,
  # select data and aesthetics
  data = c(2), color = id1, panel = analysis,
  # zoom in on time interval
  time_interval = c(1000, 3000),
  # peak labels for all peaks > 2V
  peak_bounds = TRUE,
  peak_marker = FALSE,
  peak_label = iso_format(compound, d2H),
  #peak_label_size = 3,
  peak_label_filter = compound != "unknown"
) 
```

## Data

```{r}
sample_files_w_ids %>% 
  iso_get_peak_table(include_file_info = everything()) %>% 
  filter(compound != "unknown") %>% 
  mutate(sample_label = iso_format(control, inc_time, with_glu)) %>% 
  iso_plot_data(
    x = compound, y = c(area2, d2H), color = sample_label,
    points = TRUE
  )
```

